\begin{thebibliography}{10}

    \Bibitem{modest-rht}
    \by M.\,F.~Modest
    \book Radiative Heat Transfer
    \publ Academic Press
    \yr 2003

    \Bibitem{OControl-3}
    \vol 33
    \issue 2
    \pages 157--175
    \yr 2012
    \by Clever D. and Lang J.
    \paper Optimal control of radiative heat transfer in glass cooling with restrictions on the temperature gradient

    \Bibitem{tse-lasor}
    \by O.~Tse, R.~Pinnau, N.~Siedow
    \paper Identification of temperature dependent parameters in laser--interstitial thermo therapy
    \jour Math. Models Methods Appl. Sci.
    \vol 22
    \issue 9
    \yr 2012
    \pages 1--29O

    \Bibitem{pinnau-identification}
    \by N.~Siedow O.~Tse, R.~Pinnau.
    \paper Identification of temperature dependent parameters in a simplified radiative heat transfer
    \jour Aust. J. Basic Appl. Sci.
    \pages 7--14
    \yr 2011

    \Bibitem{pinnau-optimal-control}
    \by R.~Pinnau O.~Tse
    \paper Optimal control of a simplified natural convection-radiation model
    \jour Commun. Math. Sci.
    \pages 679--707
    \yr 2013

    \Bibitem{pinnau-glass}
    \by Thomes G., Pinnau R., Seaid M., Gotz T., and A.~Klar.
    \pages Numerical methods and optimal control for glass cooling processes
    \jour Trans. Theory Stat Phys.
    \vol 31
    \issue 4--6
    \pages 513–529
    \yr 2002


    \Bibitem{covt-last}
    \by {Alexander Yu.} Chebotarev, {Andrey E.} Kovtanyuk, {Gleb V.} Grenkin, {Nikolai D.} Botkin, and {Karl Heinz} Hoffmann
    \paper Nondegeneracy of optimality conditions in control problems for a radiative-conductive heat transfer model
    \jour Applied Mathematics and Computation
    \vol 289
    \pages 371--380
    \issue 10
    \yr 2016

    \Bibitem{grenkin-13}
    Chebotarev A., Kovtanyuk A., Grenkin G., Botkin N., and Hoffman K.-H.
    \paper Boundary optimal control problem of complex heat transfer model
    \jour J. Math. Anal. Appl.
    \vol 433
    \issue 2
    \pages 1243–1260
    \yr 2016

    \Bibitem{grenkin-15}
    \by K.~Glashoff and E.~Sachs
    \paper On theoretical and numerical aspects of the bang-bang-principle
    \jour Numer. Math.
    \vol 29
    \issue 1
    \pages 93–113
    \yr 1977

    \Bibitem{cheb-origin}
    \by Kovtanyuk Andrey~E., Chebotarev Alexander~Yu., Botkin Nikolai~D., and Hoffmann Karl-Heinz
    \paper Theoretical analysis of an optimal control problem of conductive convective radiative heat transfer
    \jour J. Math. Anal. Appl.
    \vol 412
    \yr 2014
    \pages 520–528

    \RBibitem{grenkin-optimalnoe-upravleine}
    \by Гренкин Г.В.
    \paper  Оптимальное управление в нестационарной модели сложного теплообмена
    \jour Дальневост. матем. журн.
    \vol 14
    \issue 2
    \yr 2014
    \pages 160–172

    \RBibitem{a6}
    \by Р.~В.~Бризицкий, Ж.~Ю.~Сарицкая
    \paper Устойчивость решений экстремальных задач для нелинейного уравнения конвекции–диффузии–реакции при условии Дирихле
    \jour Ж. вычисл. матем. и матем. физ.
    \yr 2016
    \vol 56
    \issue 12
    \pages 2042--2053

    \RBibitem{a7}
    \by Г.~В.~Алексеев, Р.~В.~Бризицкий, Ж.~Ю.~Сарицкая
    \paper Оценки устойчивости решений экстремальных задач для нелинейного уравнения конвекции-диффузии-реакции
    \jour Сиб. журн. индустр. матем.
    \yr 2016
    \vol 19
    \issue 2
    \pages 3--16

    \Bibitem{OControl-1}
    \by Pinnau R.
    \paper Analysis of optimal boundary control for radiative heat transfer modeled by the $sp_1$-system
    \jour Comm. Math. Sci.,
    \vol 5
    \issue 4
    \pages 951–969
    \yr 2007

    \Bibitem{lemma-proof}
    \by Kovtanyuk A.E., Chebotarev A.Yu., Botkin N.D., and Hoffman Karl-Heinz
    \paper Unique solvability of a steady-state complex heat transfer model,
    \jour Commun. Nonlinear Sci. Numer. Simulat.
    \vol 20
    \pages 776–784
    \yr  2015

    \Bibitem{theorem-proof-18}
    \by A.D. Ioffe and V.M. Tikhomirov
    \book Theory of extremal problems
    \publaddr North Holland, Amsterdam
    \yr 1979

\end{thebibliography}

